---
title: "Modern Data Mining - HW 3"
author:
- Group Member 1
- Group Member 2
- Group Member 3
output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height=4, fig.width=6, warning = F)

# constants for homework assignments
hw_num <- 3
hw_due_date <- "17, March, 2019"



pacman::p_load(tidyverse, nycflights13, gapminder, ggthemes, data.table, ISLR, leaps, car, GGally, reshape2, glmnet, mapproj, bestglm, pROC, caret)
```



## Overview / Instructions

This is homework #`r paste(hw_num)` of STAT 471/571/701. It will be **due on `r paste(hw_due_date)` by 11:59 PM** on Canvas. You can directly edit this file to add your answers. Submit the Rmd file, a PDF or word or HTML version with **only 1 submission** per HW team. No zip files please.

**Note:** To minimize your work and errors, we provide this Rmd file to guide you in the process of building your final report. To that end, we've included code to load the necessary data files. Make sure that the following files are in the same folder as this R Markdown file:

* `FRAMINGHAM.dat`
* `Bills.subset.csv`
* `Bills.subset.test.csv`

The data should load properly if you are working in Rstudio, *without needing to change your working directory*.



## R Markdown / Knitr tips

You should think of this R Markdown file as generating a polished report, one that you would be happy to show other people (or your boss). There shouldn't be any extraneous output; all graphs and code run should clearly have a reason to be run. That means that any output in the final file should have explanations.

A few tips:

* Keep each chunk to only output one thing! In R, if you're not doing an assignment (with the `<-` operator), it's probably going to print something.
* If you don't want to print the R code you wrote (but want to run it, and want to show the results), use a chunk declaration like this: `{r, echo=F}`
* If you don't want to show the results of the R code or the original code, use a chunk declaration like: `{r, include=F}`
* If you don't want to show the results, but show the original code, use a chunk declaration like: `{r, results='hide'}`.
* If you don't want to run the R code at all use `{r, eval = F}`.
* We show a few examples of these options in the below example code. 
* For more details about these R Markdown options, see the [documentation](http://yihui.name/knitr/options/).
* Delete the instructions and this R Markdown section, since they're not part of your overall report.

## Problem 0

Review the code and concepts covered during lecture, in particular, logistic regression and classification. 

## Problem 1
We will continue to use the Framingham Data (`Framingham.dat`) so that you are already familiar with the data and the variables. All the results are obtained through training data.

To keep our answers consistent, use a subset of the data, and exclude anyone with a missing entry. For your convenience, we've loaded it here together with a brief summary about the data.

```{r data preparation, include=F}
# Notice that we hide the code and the results here
# Using `include=F` in the chunk declaration. 
hd_data <- read.csv("Framingham.dat")
str(hd_data) 

### Renames, setting the variables with correct natures...
names(hd_data)[1] <- "HD"
hd_data$HD <- as.factor(hd_data$HD)
hd_data$SEX <- as.factor(hd_data$SEX)
str(hd_data)
#tail(hd_data, 1)    # The last row is for prediction
hd_data.new <- hd_data[1407,] # The female whose HD will be predicted.
hd_data <- hd_data[-1407,]  # take out the last row 
hd_data.f <- na.omit(hd_data)
```

We note that this dataset contains 311 people diagnosed with heart disease and 1095 without heart disease.
```{r table heart disease, echo = F, comment = " "}
# we use echo = F to avoid showing this R code
table(hd_data$HD) # HD: 311 of "0" and 1095 "1" 
```

After a quick cleaning up here is a summary about the data:
```{r data summary, comment="     "}
# using the comment="     ", we get rid of the ## in the output.
summary(hd_data.f)
```
### Part 1A
Conceptual questions to understand the building blocks of logistic regression. All the codes in this part should be hidden. We will use a small subset to run a logistic regression of `HD` vs. `SBP`. 

i. Take a random subsample of size 5 from `hd_data_f` which only includes `HD` and `SBP`. Also set   `set.seed(50)`. List the five observations neatly below. No code should be shown here.
```{r, results='hide'}
N<-length(hd_data.f$HD)
N
```
```{r, results='hide'}
set.seed(10)
```

Take a random sample of n=5 from 1 to N=1393
```{r, results='hide'}
index.test <- sample(N,5)
index.test
```
Set test and training data
```{r, results='hide'}
data.test <- hd_data.f[index.test,c("HD","SBP")]  #dim(data.test)
data.test
```


ii. Write down the likelihood function using the five observations above.

\[\begin{split}
\mathcal{Lik}(\beta_0, \beta_1 \vert {\text Data}) &= {\text {Prob(the outcome of the data)}}\\
&=Prob((HD=0|SBP=150), (Y=1|SBP=124), (Y=1|SBP=130),(Y=0|SBP=176),(Y=1|SBP=138)) \\
&=Prob((HD=0|SBP=150) \times Prob(Y=1|SBP=124) \times Prob(Y=1|SBP=130) \times Prob(Y=0|SBP=176)\times Prob(Y=1|SBP=138) \\
&= \frac{1}{1+e^{\beta_0 + 150 \beta_1}}\cdot\frac{1}{1+e^{\beta_0 + 176\beta_1}}\cdot\frac{e^{\beta_0 + 124 \beta_1}}{1 + e^{\beta_0 + 124 \beta_1}} \cdot\frac{e^{\beta_0 + 130 \beta_1}}{1 + e^{\beta_0 + 130 \beta_1}}\cdot\frac{e^{\beta_0 + 138 \beta_1}}{1 + e^{\beta_0 + 138 \beta_1}}
	\end{split}\]


iii. Find the MLE based on this subset using glm(). Report the estimated logit function of `SBP` and the probability of `HD`=1. Briefly explain how the MLE are obtained based on ii. above.
```{r, results='hide'}
fit1 <- glm(HD~SBP, data.test, family=binomial(logit)) 
```

```{r, results='hide'}
summary(fit1, results=TRUE)
names(fit1)
```


To see the prob function estimated by glm: 

- logit = -551.38 - 3.83 SBP

- $P(HD = 1 \vert SBP) = \frac{e^{-551.38 - 3.83 \times  SBP}}{1+e^{-551.38 - 3.83 \times SBP}}$

- $P(HD = 0 \vert SBP) = \frac{1}{1+e^{-551.38 - 3.83 \times SBP}}$

We used the 5 know observation to build a likelihood function, The maximizes $\hat \beta_1$ and $\hat \beta_0$ will be used to estimate the unknown parameters


### Part 1B 


Goal: Identify important risk factors for `Heart.Disease.` through logistic regression. 
Start a fit with just one factor, `SBP`, and call it `fit1`. Let us add one variable to this at a time from among the rest of the variables. 
```{r, results='hide'}
fit1 <- glm(HD~SBP, hd_data.f, family=binomial)
summary(fit1)
fit1.1 <- glm(HD~SBP + AGE, hd_data.f, family=binomial)
summary(fit1.1)
fit1.2 <- glm(HD~SBP + SEX, hd_data.f, family=binomial)
summary(fit1.2)
fit1.3 <- glm(HD~SBP + DBP, hd_data.f, family=binomial)
summary(fit1.3)
fit1.4 <- glm(HD~SBP + CHOL, hd_data.f, family=binomial)
summary(fit1.4)
fit1.5 <- glm(HD~SBP + DBP, hd_data.f, family=binomial)
summary(fit1.5)
fit1.6 <- glm(HD~SBP + FRW, hd_data.f, family=binomial)
summary(fit1.6)
fit1.7 <- glm(HD~SBP + CIG, hd_data.f, family=binomial)
summary(fit1.7)
```





i. Which single variable would be the most important to add? Add it to your model, and call the new fit `fit2`.  

We will pick up the variable either with highest $|z|$ value, or smallest $p$ value. From all the two variable models we see that `SEX` will be the most important addition on top of the SBP. And here is the summary report.
```{r the most important addition, results='asis', comment="   "}
## How to control the summary(fit2) output to cut some junk?
## We could use packages: xtable or broom. 
library(xtable)
options(xtable.comment = FALSE)
fit2 <- glm(HD~SBP + SEX, hd_data.f, family=binomial)
xtable(fit2)
```
ii. Is the residual deviance of `fit2` always smaller than that of `fit1`? Why or why not?
  
  Yes. the smaller deviance means the better the model expains the data. Just like R2, the more variables a model has, the smaller error we have. Thus, fit2 always has smaller errors than fit1, which means the residual deviance of fit2 is always smaller than that of fit1


  
iii. Perform both the Wald test and the Likelihood ratio tests (Chi-Squared) to see if the added variable is significant at the .01 level.  What are the p-values from each test? Are they the same? 

The Wald Test, 
since the p-value for sex is 1.02e-10 which is much smaller than .05. We reject the null hypothesis to say sex is significant at 0.05 level.
```{r, results='hide'}
summary(fit2)
confint.default(fit2) 
```

Likelihood Ratio Test.
Bases on likelihood ratio test, we find the p-value is 1.829947e-21 (3.828e-11 in annova test), it means sex is significant at 0.05 level, and we reject null hypothesis

- Null deviance = 1469.3
- Residual deviance: 1373.8  
- $\chi^2 = 1469.3-1373.8= 95.5$

Get the $\chi^2$ statistic
```{r, results='hide'}
chi.sq <- 1469.3-1373.8
```
$p_{value}$: from the likelihood Ratio test
```{r, results='hide'}
pchisq(chi.sq, 2, lower.tail=FALSE)

```

```{r, results='hide'}
anova(fit2, test="Chisq")
```


### Part 1C -  Model building

Start with all variables. Our goal is to fit a well-fitting model, that is still small and easy to interpret (parsimonious).

i. Use backward selection method. Only keep variables whose coefficients are significantly different from 0 at .05 level. Kick out the variable with the largest p-value first, and then re-fit the model to see if there are other variables you want to kick out.

```{r, results='hide'}
fit3 <- glm(HD~., hd_data.f, family=binomial)
summary(fit3)
```
Kick out DBP with largest p-value
```{r, results='hide'}
fit3.1 <-update(fit3,.~. -DBP)
summary(fit3.1)
```
Kick out FRW with largest p-value.
```{r, results='hide'}
fit3.2 <-update(fit3.1,.~. -FRW)
summary(fit3.2)
```
CIG didn't really add much to decrease the residual deviance
```{r, results='hide'}
fit3.3 <- update(fit3.2, .~. -CIG)
summary(fit3.3)
```
All variables are significant now. Stop here. 


ii. Use AIC as the criterion for model selection. Find a model with small AIC through exhaustive search. Does exhaustive search  guarantee that the p-values for all the remaining variables are less than .05? Is our final model here the same as the model from backwards elimination? 

Get the design matrix without 1's and HD.
```{r, results='hide'}
Xy <- model.matrix(HD ~.+0, hd_data.f) 
```

Attach y as the last column.
```{r, results='hide'}
Xy <- data.frame(Xy, hd_data.f$HD)   
```

```{r, results='hide'}
str(Xy)
```

```{r, results='hide'}
fit.all <- bestglm(Xy, family = binomial, method = "exhaustive", IC="AIC", nvmax = 10)
names(fit.all) 
```

List the top 5 models. In the way any one of the following model could be used. 
```{r, results='hide'}
fit.all$BestModels  
```


```{r, results='hide'}
summary(fit.all$BestModel)
```


```{r, results='hide'}
fit4 <- glm(HD~ AGE+SEX+SBP+CHOL+FRW+CIG, hd_data.f, family=binomial)
summary(fit4)
```
Compare to Backward selection, Lowest AIC selection has more variables and the model does not guarantee all variables are significant at 0.05 level.



iii. Use the model chosen from part ii. as the final model. Write a brief summary to describe important factors relating to Heart Diseases (i.e. the relationships between those variables in the model and heart disease). Give a definition of “important factors”. 

Based on model from part ii, we found that Age, being Male, SBP, CHOL, FRW and CIG are the factors relating to Heart Diseases.
- Holding everything else equal, the log odds ratio of getting Heart Disease will increase by 0.061 for 1 unit increase in Age
- Holding everything else equal, the log odds ratio of getting Heart Disease will increase by 0.911 for a person being male compared to a female.
- Holding everything else equal, the log odds ratio of getting Heart Disease will increase by 0.015 for 1 unit increase in SBP
- Holding everything else equal, the log odds ratio of getting Heart Disease will increase by 0.0044 for 1 unit increase in CHOL.
- Holding everything else equal, the log odds ratio of getting Heart Disease will increase by 0.006 for 1 unit increase in FRW
- Holding everything else equal, the log odds ratio of getting Heart Disease will increase by 0.012 for 1 unit increase in CIG.


### Part 1D - Prediction
Liz is a patient with the following readings: `AGE=50, GENDER=FEMALE, SBP=110, DBP=80, CHOL=180, FRW=105, CIG=0`. What is the probability that she will have heart disease, according to our final model?


The probability for Liz to have heart disease is 0.0496.

```{r}
hd_data.liz <- data.frame("HD"=NA, 
                          "AGE"=50, 
                          "SEX"="FEMALE", 
                          "SBP"=110, 
                          "DBP"=80,
                          "CHOL"=180,
                          "FRW"=105,
                          "CIG"=0)
predict.fit.AIC.f <- predict(fit4, hd_data.liz, type="response")
predict.fit.AIC.f 

```








### Part 2 - Classification analysis

a. Display the ROC curve using `fit1`. Explain what ROC reports and how to use the graph. Specify the classifier such that the False Positive rate is less than .1 and the True Positive rate is as high as possible.



The ROC curve shows the tradeoff between specificity and sensitivity for different thresholds The overall performance of the classifier is given by the area under the curve (AUC). The larger the AUC, the better the classifier performs. An ideal ROC curve will hug the top left corner and will have an AUC equal to or very close to 1. For our case, the AUC is 0.6,and the plotted ROC curve is relatively flat.

The max sensitivity of which the False Positive rate is less than 0.1 is 0.2149837, and the threshold is 0.2982114.

```{r}
fit1.roc <- roc(hd_data.f$HD,fit1$fitted.values, plot=TRUE)

```
```{r}
pROC::auc(fit1.roc) 
```

```{r}
plot(1-fit1.roc$specificities, fit1.roc$sensitivities, col="red", pch=16,
     xlab="False Positive", 
     ylab="True Positive")
```

```{r}
fit1.roc.result <- data.frame(fit1.roc$specificities, fit1.roc$sensitivities, fit1.roc$thresholds)
fit1.roc.result %>% filter(fit1.roc.specificities>=0.9) %>% filter(fit1.roc.sensitivities==max(fit1.roc.sensitivities))
```




b. Overlay two ROC curves: one from `fit1`, the other from `fit2`. Does one curve always contain the other curve? Is the AUC of one curve always larger than the AUC of the other one? Why or why not?


The two curves don't always contain each other. The blue line rises higher than the red line when false positive is greater than 0.2.
AUC of fit 2 is always larger than AUC of fit 1. AUC measures the performance of how well the classifer as a whole predict the data. Thus the higher AUC means the better classifer. In this case, fit2 has more predictors with significant p-value (SEX), thus have smaller errors and better predicting power.

```{r}
fit2.roc <- roc(hd_data.f$HD, fit2$fitted, plot=F, col="blue")
plot(1-fit1.roc$specificities, fit1.roc$sensitivities, col="red", pch=16,
     xlab=paste("False Positive
                ","AUC(fit1) =",
                round(pROC::auc(fit1.roc),2),
                "  AUC(fit2) =",
                round(pROC::auc(fit2.roc),2) ), 
     ylab="Sensitivities")

points(1-fit2.roc$specificities, fit2.roc$sensitivities, col="blue", pch=16)
legend("topleft", legend=c("fit1", "fit2"),
       lty=c(1,1), lwd=c(2,2), col=c("red", "blue"))
```




c. Estimate the Positive Prediction Values and Negative Prediction Values for `fit1` and `fit2` using .5 as a threshold. Which model is more desirable if we prioritize the Positive Prediction values?

Positive prediction of fit1 is 0.45, and that of fit is 0.47, so fit2 is more desirable. 

Fit 1
```{r}
fit1.pred.5 <- ifelse(fit1$fitted.values > 0.5, "1", "0")
cm.1 <- table(fit1.pred.5, hd_data.f$HD)
cm.1
```
Fit 1, Positive prediction
```{r}
positive.pred.1 <- cm.1[2, 2] / (cm.1[2, 1] + cm.1[2, 2])
positive.pred.1
```
Fit1 Negative prediction
```{r}
negative.pred.1 <- cm.1[1,1]/(cm.1[1,2]+cm.1[1,1])
negative.pred.1
```


Fit2

```{r}
fit2.pred.5 <- ifelse(fit2$fitted.values > 0.5, "1", "0")
cm.2 <- table(fit2.pred.5, hd_data.f$HD)
cm.2
```

Positive prediction
```{r}
positive.pred.2 <- cm.2[2, 2] / (cm.2[2, 1] + cm.2[2, 2])
positive.pred.2
```

Negative prediction
```{r}
negative.pred.2 <- cm.2[1,1]/(cm.2[1,2]+cm.2[1,1])
negative.pred.2
```



d. (Optional/extra credit) For `fit1`: overlay two curves,  but put the threshold over the probability function as the x-axis and positive prediction values and the negative prediction values as the y-axis.  Overlay the same plot for `fit2`. Which model would you choose if the set of positive and negative prediction values are the concerns? If you can find an R package to do so, you may use it directly.

 

  
  
  
  
### Part 3 - Bayes Rule
Bayes rules with risk ratio $\frac{a_{10}}{a_{01}}=10$ or $\frac{a_{10}}{a_{01}}=1$. Use your final model obtained from 1 B) to build a class of linear classifiers.


a. Write down the linear boundary for the Bayes classifier if the risk ratio of $a_{10}/a_{01}=10$.
$a_{01}/a_{10}=1/10=0.1$
$P(Y=1 \vert x) > \frac{\frac{a_{0,1}}{a_{1,0}}}{1 + \frac{a_{0,1}}{a_{1,0}}}$
$P(Y=1 \vert x) > \frac{0.1}{1 + 0.1}=0.09$
$logit > log\frac{0.09}{0.91} = -2.31$
So the linear boundry is 
-9.2279 (Intercept) + 0.0615 AGE + 0.9113 SEXMALE + 0.016 SBP + 0.0045 CHOL + 0.006 FRW + 0.0123 CIG  >=  -2.31
```{r}
coef.fit <- paste(round(coef(fit4),4), names(coef(fit4)))
print(paste(paste(coef.fit, collapse=" + " ), " >= ", -2.31))
```



b. What is your estimated weighted misclassification error for this given risk ratio?

MCE=0.716

```{r}
fit.bayes <- as.factor(ifelse(fit4$fitted.values>0.09,"1","0"))
```

```{r}
MCE.bayes=(sum(10*(fit.bayes[hd_data.f$HD == "1"] != "1")) 
           + sum(fit.bayes[hd_data.f$HD == "0"] != "0"))/length(hd_data.f$HD)
MCE.bayes
```


c. Recall Liz, our patient from part 1. How would you classify her under this classifier?

The probability for Liz to have heart disease is 0.0496, which is under thr threshold of 0.09, so Liz will be classified as "0" -no heart disease.


Now, draw two estimated curves where x = posterior threshold, and y = misclassification errors, corresponding to the thresholding rule given in x-axis.

```{r}
fit4.roc <- roc(hd_data.f$HD, fit4$fitted.values, plot = F)

plot(fit4.roc$thresholds, 1-fit4.roc$specificities + 1 - fit4.roc$sensitivities,  col="green", pch=16,  
     xlab="Threshold",
     ylab="MCE",
     main = "Thresholds vs. MCE")
```

In this graph, we are seeing MCE changes in reponse to different thresholds. It's true that when threshold is 0 or 1, misclassification error is 1 because these thresholds are too extreme, classifying the whole data to be either "0" or "1". The curve dips at threshold = 0.2, which means classification error is minimized at this threshold. And then it starts going up as threshold gets larger. We interpret this as trade-off between sensitivity and specifities happening, which is expressed clearer when we examines sensitivities and specificities separately as below.

```{r}
plot(fit4.roc$thresholds, fit4.roc$sensitivities,  col="green", pch=16,  
     xlab="Threshold",
     ylab="Sensitivities",
     main = "Thresholds vs. Sensitivities")
```

```{r}
plot(fit4.roc$thresholds, 1- fit4.roc$specificities,  col="green", pch=16,  
     xlab="Threshold",
     ylab="False Positive",
     main = "Thresholds vs. False Positive")
```

```{r}
plot(fit4.roc$thresholds, fit4.roc$specificities,  col="green", pch=16,  
     xlab="Threshold",
     ylab="Specificities",
     main = "Thresholds vs. Specificities")
```






d. Use weighted misclassification error, and set $a_{10}/a_{01}=10$. How well does the Bayes rule classifier perform? 

MCE for this classifier is 0.716 which is relatively high, which means the performence of this classifier is relatively poor. 

we can also check its confusion matrix. The sensitivity of this classifier is 0.97 which is high. It means we identified most true positive. However, the specificity is 0.16, means we have a lot of false positive included. The accuracy is 0.34 which also indicate the performance of this classifier is poor.
```{r}
confusionMatrix(data = fit.bayes, reference = hd_data.f$HD, positive = levels(hd_data.f$HD)[2])
```



e. Use weighted misclassification error, and set $a_{10}/a_{01}=1$. How well does the Bayes rule classifier perform? 

The MCE of this classifier is 0.217, which indicate a better performence. We also checked the confusion matrics. In this classifier, we have a lower sensitivity (0.088), which means this classifier is not very good at identifying tru postitive values (people who actually have heart disease). However, we have a very high specificity (0.98), which means this classifier is very good at identifying people who does not have a heart disease 

$a_{01}/a_{10}=1/10=1$
$P(Y=1 \vert x) > \frac{\frac{a_{0,1}}{a_{1,0}}}{1 + \frac{a_{0,1}}{a_{1,0}}}$
$P(Y=1 \vert x) > \frac{1}{1 + 1}=0.5$

```{r}
fit1.bayes <- as.factor(ifelse(fit4$fitted.values>0.5,"1","0"))
```

```{r}
MCE.bayes1=(sum((fit1.bayes[hd_data.f$HD == "1"] != "1")) 
           + sum(fit1.bayes[hd_data.f$HD == "0"] != "0"))/length(hd_data.f$HD)
MCE.bayes1
```

```{r}
confusionMatrix(data = fit1.bayes, reference = hd_data.f$HD, positive = levels(hd_data.f$HD)[2])
```







## Problem 2

How well can we predict whether a bill will be passed by the legislature? 

Hundreds to thousands of bills are written each year in Pennsylvania. Some are long, others are short. Most of the bills do not even get to be voted on (“sent to the floor”). The chamber meets for 2-year sessions.  Bills that are not voted on before the end of the session (or which are voted on but lose the vote) are declared dead. Most bills die. In this study we examine about 8000 bills proposed since 2009, with the goal of building a classifier which has decent power to forecast which bills are likely to be passed. 

We have available some information about 8011 bills pertaining to legislation introduced into the Pennsylvania House of Representatives.  The goal is to predict which proposals will pass the House. Here is some information about the data:

The response is the variable called `status.` `Bill:passed` means that the bill passed the House; `governor:signed` means that the bill passed both chambers (including the House) and was enacted into law; `governor:received` means that the bill has passed both chambers and was placed before the governor for consideration.  All three of these statuses signify a success or a PASS (Meaning that the legislature passed the bill. This does not require it becoming law). All other outcomes are failures.

Here are the rest of the columns:

*	`Session` – in which legislative session was the bill introduced
*	`Sponsor_party` – the party of the legislator who sponsored the bill (every bill has a sponsor)
*	`Bill_id` – of the form HB-[bill number]-[session], e.g., `HB-2661-2013-2014` for the 2661st House Bill introduced in the 2013-2014 session.
*	`Num_cosponsors` – how many legislators cosponsored the bill
*	`Num_d_cosponsors` – how many Democrats cosponsored the bill
*	`Num_r_cosponsors` – how many Republicans cosponsored the bill
*	`Title_word_count` – how many words are in the bill’s title
*	`Originating_committee` – most bills are sent (“referred”) to a committee of jurisdiction (like the transportation committee, banking & insurance committee, agriculture & rural affairs committee) where they are discussed and amended.  The originating committee is the committee to which a bill is referred.
*	`Day_of_week_introduced` – on what day the bill was introduced in the House (1 is Monday)
*	`Num_amendments` – how many amendments the bill has
*	`Is_sponsor_in_leadership` – does the sponsor of the bill hold a position inside the House (such as speaker, majority leader, etc.)
*	`num_originating_committee_cosponsors` – how many cosponsors sit on the committee to which the bill is referred
*	`num_originating_committee_cosponsors_r` – how many Republican cosponsors sit on the committee to which the bill is referred
*	`num_originating_committee_cosponsors_d` - how many Democratic cosponsors sit on the committee to which the bill is referred

The data you can use to build the classifier is called `Bills.subset`. It contains 7011 records from the full data set. I took a random sample of 1000 bills from the 2013-2014 session as testing data set in order to test the quality of your classifier, it is called `Bills.subset.test.`

Your job is to choose a best set of classifiers such that

* The testing ROC curve pushes to the upper left corner the most, and has a competitive AUC value.
* Propose a reasonable loss function, and report the Bayes rule together with its weighted MIC. 
* You may also create some sensible variables based on the predictors or make other transformations to improve the performance of your classifier.

Here is what you need to report: 

1. Write a summary about the goal of the project. Give some background information. If desired, you may go online to find out more information.
2. Give a preliminary summary of the data. 
3. Based on the data available to you, you need to build a classifier. Provide the following information:
    *	The process of building your classifier
    *	Methods explored, and why you chose your final model
    *	Did you use a training and test set to build your classifier using the training data? If so, describe the process including information about the size of your training and test sets.
    *	What is the criterion being used to build your classifier?
    *	How do you estimate the quality of your classifier?
4. Suggestions you may have: what important features should have been collected which would have helped us to improve the quality of the classifiers.

*Final notes*: The data is graciously lent from a friend. It is only meant for you to use in this class. All other uses are prohibited without permission. 


Give a preliminary summary of the data. 
```{r}
bill_data <- read.csv("Bills.subset.csv")
bill_data_test <-read.csv("Bills.subset.test.csv")
```


```{r}
dim(bill_data)
str(bill_data)
```
```{r}
dim(bill_data_test)
```

```{r}
summary(bill_data)
```

```{r}
sapply(bill_data, function(x) sum(is.na(x)))  #find out missing values in training data
```
```{r}
sapply(bill_data_test, function(x) sum(is.na(x)))
```

```{r}
#replace all missing day of week with random number from 1 to 6 (Mon to Sat)
set.seed(23)
bill_data[is.na(bill_data[,"day.of.week.introduced"]),"day.of.week.introduced"] <- sample(1:6,1)  
bill_data_test[is.na(bill_data_test[,"day.of.week.introduced"]),"day.of.week.introduced"] <- sample(1:6,1)  

```




```{r}
#deselect unrelevant column "bill_id"
bill_data <- bill_data[,-1] 
bill_data_test <- bill_data_test[,-1]
```

```{r}
dim(bill_data)
dim(bill_data_test)
```

```{r}
# We create a column to identify 3 status "bill:passed", "governor:signed", "governor:received", and differ from other status
bill_data <- bill_data %>% mutate(passed = status %in% c("bill:passed", "governor:signed", "governor:received"))
bill_data_test <- bill_data_test %>% mutate(passed = status %in% c("bill:passed", "governor:signed", "governor:received"))
```

```{r}
str(bill_data_test)
```



```{r}
#turn categorical data into factors
bill_data$passed <- as.factor(as.numeric(bill_data$passed))
bill_data$is_sponsor_in_leadership = as.factor(bill_data$is_sponsor_in_leadership)
bill_data$session <- as.factor(bill_data$session)
bill_data$sponsor_party <- as.factor(bill_data$sponsor_party)
bill_data$originating_committee <- as.factor(bill_data$originating_committee)
bill_data$day.of.week.introduced <- as.factor(bill_data$day.of.week.introduced)


bill_data_test$passed <- as.factor(as.numeric(bill_data_test$passed))
bill_data_test$is_sponsor_in_leadership = as.factor(bill_data_test$is_sponsor_in_leadership)
bill_data_test$session <- as.factor(bill_data_test$session)
bill_data_test$sponsor_party <- as.factor(bill_data_test$sponsor_party)
bill_data_test$originating_committee <- as.factor(bill_data_test$originating_committee)
bill_data_test$day.of.week.introduced <- as.factor(bill_data_test$day.of.week.introduced)


```

```{r}
summary(bill_data)
```








